Effects of Antenna Beam Chromaticity on Redshifted 21~cm Power Spectrum and Implications for Hydrogen Epoch of Reionization Array

### 2016-03-06 ForegroundSimulations/ms.tex
Update sensitivity figures, minor text edits
From ed32dfb (fgsim):
====================

(*) Update sensitivity figures

In ms.tex,

(*) Minor edits throughout
(*) Use updated sensitivity figures

\shorttitle{Effects of Beam Chromaticity on EoR Power Spectra Measurements}
\shortauthors{Thyagarajan et~al.}

@@

Unaccounted systematics from foregrounds and instruments prevent present day experiments from detecting redshifted 21~cm signal from the Epoch of Reionization (EoR). In-depth understanding of these effects is critical for the success of such experiments. We identify spectral structures (chromaticity) in the antenna power pattern as one such systematic that significantly affects all EoR measurements. Increase in chromaticity of the antenna beam causes the already bright foreground power to leak well beyond ideal limits and contaminate the otherwise clean EoR signal modes. A primary source of this chromaticity is reflections in the antenna-feed assembly and between structures in neighboring antennas. Using precise foreground simulations taking wide-field effects into account, we provide a framework to set cosmologically-motivated upper limits on these reflections to prevent further EoR signal degradation. We use the Hydrogen Epoch of Reionization Array (HERA) as an example. Using a simple and yet high dynamic range spectral weighting scheme, we demonstrate that EoR detection using HERA will not be impeded by such spectral systematics even in a conservative scenario that does not require removal of foregrounds. We also show that all baselines in a 19-element HERA layout are capable of detecting EoR over a substantial observing window on the sky.

@@

The period in the history of the Universe characterized by the transition of neutral hydrogen in the intergalactic medium (IGM) to a fully ionized state due to the formation of radiating objects such as the first stars and galaxies is referred to as the Epoch of Reionization (EoR). This is an important period of nonlinear growth of matter density perturbations and astrophysical evolution leading to the large scale structure observed currently in the Universe. And yet, this period in the Universe's history has remained poorly probed to date with observations.

The redshifted neutral hydrogen from the IGM in this epoch has been identified to be one of the most promising and direct probes of the EoR \citep{sun72,sco90,mad97,toz00,ili02}. Numerous experiments using low frequency radio telescopes targeting the redshifted 21~cm line from the spin-flip transition of H{\sc i} have become operational such as the Murchison Widefield Array \citep[MWA;][]{lon09,bow13,tin13}, the Donald~C.~Backer Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10}, the Low Frequency Array \citep[LOFAR;][]{van13} and the Giant Metrewave Radio Telescope EoR experiment \citep[GMRT;][]{pac13}. These instruments have sufficient sensitivity for a statistical detection of the EoR signal via estimating the spatial power spectrum of the redshifted H{\sc i} temperature fluctuations \citep{bea13,thy13}. These instruments are intended to be precursors and pathfinders to the next generation of low frequency radio observatories such as the Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} \citep[HERA;][]{deb16} and the Square Kilometre Array\footnote{\url{https://www.skatelescope.org/}} (SKA). These next-generation instruments will advance the capability from a mere statistical detection of the signal to a direct three-dimensional tomographic imaging of the H{\sc i} from the EoR.

@@

When expressed in the coordinate system of power spectrum measurements described by the three-dimensional wavenumber ($k$), the foreground emission is restricted to a wedge-shaped region commonly referred to as the {\it foreground wedge} \citep{bow09,liu09,liu14a,liu14b,dat10,liu11,gho12,mor12,par12b,tro12,ved12,dil13,pob13,thy13,dil14} whereas the EoR signal has spherical symmetry due to its isotropy which appears elongated along line of sight $k$ modes due to peculiar velocity effects when dominated by matter density perturbations during early stages of reionization.

However, despite attempts to exploit this knowledge about differences in spectral and spatial symmetry in lengthy observations, analysis from the first-generation experiments has at best only led to weak constraints on reionization models \citep{ali15,dil15,pob15}. This is because their sensitivity is primarily limited by instrumental and foreground systematics. The extreme dynamic range required to suppress foreground and instrument systematics demands high precision modeling of foregrounds as observed by modern wide-field instruments. One such study was undertaken by \citet{thy15a,thy15b}. Their studies of effects of wide-field measurements of EoR power spectra have demonstrated the {\it pitchfork} effect wherein foreshortening of baselines causes a prominent enhancement of foreground power near the horizon limits of the {\it foreground wedge} as well as in the contamination beyond. It highlighted the importance of controlling the response of wide-field instruments near the horizon. The need of precise subtraction of foregrounds in wide-field measurements was also highlighted in \citet{pob16}.

In this paper, we explore yet another phenomenon that inherently extends foreground power beyond the {\it wedge}, namely, that arising from the chromaticity of the antenna power pattern. While highlighting the significance of such instrumental systematics on future EoR experiments, we demonstrate that the current HERA instrument design will exceed these specifications to detect the EoR signal with high significance, when data is limited by foregrounds and systematics, even in the most conservative scenario which involves neither foreground subtraction nor optimal analysis techniques but only relies on a simple spectral weighting scheme.

This paper is organized as follows. \S\ref{sec:HERA} introduces the HERA instrument. A brief summary of the delay spectrum technique used extensively in this analysis and the recently confirmed {\it pitchfork} effect are presented in \S\ref{sec:delay-spectrum}. \S\ref{sec:sim} describes foreground simulations including antenna beam pattern and all-sky foreground models. \S\ref{sec:beam-chromaticity} investigates the effects of chromaticity of antenna beam on the resulting delay power spectrum and the cosmologically motivated constraints it places on reflections in the instrument. Lower limits on sensitivity of HERA under conservative conditions are demonstrated in \S\ref{sec:eor-sensitivity}. Our findings are summarized in \S\ref{sec:summary}.

@@

The delay spectrum has a close resemblance to cosmological H~{\sc i} spatial power spectrum. Thus, the delay power spectrum is defined as:
\begin{align}\label{eqn:delay-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{1}{\Omega\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \label{eqn:kperp-baseline}\\
  k_\parallel &\equiv \frac{2\pi\tau\,f_{21}H_0\,E(z)}{c(1+z)^2}, \label{eqn:kprll-delay}
\end{align}
where, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H~{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology, and following \citet{par14},
\begin{align}
  \Omega\Delta B &= \iint \left|A(\hat{\boldsymbol{s}},f)\,W(f)\right|^2\,\dif\Omega\,\dif f.
\end{align}

@@

We note that these visibilities and delay power spectra are dependent on right ascension (RA) of the pointing as well as the frequency band. We defer expressing visibilities and delay power spectra explicitly as a function of pointing or center frequency (or redshift) to later sections where it becomes necessary.

@@

It was also demonstrated in these studies that design of antenna power pattern, specifically its amplitude near the horizon, is an important tool in mitigating foreground contamination caused by these wide-field effects. A dish characterized by a nominal {\it Airy} pattern was found to mitigate this contamination by over four and two orders of magnitude relative to a dipole and a phased array of dipoles respectively. HERA has significantly based its antenna design principles on these findings in choosing its antenna geometry while paying close attention to the properties of the resulting antenna power pattern.

In this paper, we investigate the spectral properties of the proposed dish design through their effects on the resultant foreground delay power spectrum and the constraints they place on attenuation required to suppress reflections between the dish-receiver assembly and antenna pairs to within tolerable limits.

@@

In a related series of papers, \citet{neb16} discuss the agreement of these simulated antenna beam patterns with actual measurements and \citet{ewa16} model the reflections and return loss expected in the proposed antenna-receiver assembly. Our focus in this paper is to investigate chromaticity of antenna power patterns from the point of view of their impact on foreground contamination in delay power spectra.

@@

For reference, we use two models of EoR. In the first, simulations of the H{\sc i} signal were created using the publically available 21cmFAST\footnote{\url{http://homepage.sns.it/mesinger/DexM\_\_\_21cmFAST.html}} code described in \citet{mes11}. The code uses the excursion set formalism of \citet{fur04a} to generate ionization and 21cm brightness fields for numerous redshifts. The model shown in this paper assumes the same fiducial values of $T_\text{vir}^\text{min} = 2 \times 10^4$~K (virial temperature of minimum mass of dark matter halos that host ionizing sources), $\zeta = 20$ (ionization efficiency), and $R_\text{mfp}=15$~Mpc (mean free path of UV photons) which predicts the redshift of 50\% ionization (and hence a peak in the power spectrum signal) to be at $z=8.5$ as in Ewall-Wice~et~al.(2016; submitted). The second model is borrowed from the simulations described in detail in \citet{lid08}. Hereafter, we refer to these two as EoR models 1 and 2 respectively.

@@

The equation for delay spectrum describes the mapping between sky location of a foreground object and delay. The chromatic nature (variation with frequency) of the antenna power pattern results in a convolution of the geometrical mapping with the delay response of spectral chromaticity of the power pattern. This can result in a significant spillover of foreground power beyond the horizon delay limits especially in the case of foregrounds near the horizon.

@@

Since our aim in this paper is to quantify foreground contamination and EoR sensitivity without employing any sophisticated {\it foreground removal} techniques, we turn out attention to {\it foreground avoidance} strategy that employs spectral weighting technique. In such an estimation of power spectrum, specific choices for spectral weighting function, $W(f)$, have been found to be effective in reducing foreground contamination by many orders of magnitude \citep{thy13} and is regularly used in EoR data analysis \citep{par12a}. For instance, a {\it Blackman-Harris} function \citep{har78} has a dynamic range of $\sim$100 -- 120~dB in delay power spectrum and a reduced effective bandwidth, with:
\begin{align}\label{eqn:Beff}
  \epsilon &= \frac{1}{B} \int_{-B/2}^{+B/2} |W(f)|^2\,\dif f, \\
  \textrm{and,}\quad B_\textrm{eff} &= \epsilon\,B,
  % B_\textrm{eff} &= \int_{-B/2}^{+B/2} |W(f)|^2\,\dif f.
\end{align}
where, $1-\epsilon$ is the loss in overall spectral sensitivity and $B_\textrm{eff}$ is the effective bandwidth. For a {\it Blackman-Harris} window function denoted by $W_\textrm{BH}(f)$, $\epsilon \approx 50\%$. Hence, the window spans a total of $2N$ frequency channels each of width $\Delta f$ to have $B_\textrm{eff} \approx N\,\Delta f$.

Considering that the dynamic range required to suppress foreground contamination in the {\it EoR window} may probably be even higher than that provided by a {\it Blackman-Harris} window function, we use a modified version given by the convolution of a {\it Blackman-Harris} window with itself:
\begin{align}\label{eqn:bhw2}
  W(f) &= W_\textrm{BH}(f) \ast W_\textrm{BH}(f).
\end{align}
In its Fourier domain, this window has a response obtained by squaring the {\it Blackman-Harris} window response and thus increases the dynamic range further in the power response by another 100 -- 120~dB while $\epsilon\approx 42\%$. This is an enormous gain in sidelobe characteristics for a negligible loss in over sensitivity relative to a simple {\it Blackman-Harris} weighting. This will ensure that the contamination resulting from spillover of foregrounds along the line-of-sight $k$-modes are limited by intrinsic spectral structures in the foregrounds and the beam patterns and not by sidelobes from spectral weighting. In this paper, we apply this modified spectral weighting function defined in equation~\ref{eqn:bhw2} whenever foregrounds are represented in the Fourier delay domain.

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{window_function_modifications.eps}
  \caption{Choices for spectral weighting functions, $W(f)$ ({\it top}) and their delay power spectrum responses, $|\widetilde{W}(\tau)|^2$ ({\it bottom}). The gray curves correspond to a {\it Blackman-Harris} window while those in black correspond to a {\it Blackman-Harris} window convolved with itself. The overall sensitivity of the former is $\epsilon \approx 50\%$ of a rectangular window while that of the latter is $\epsilon \approx 42\%$. Hence, the latter is narrower than the former. As a result, the main lobe of the convolved window function response is wider than that of a {\it Blackman-Harris} window. However, the sidelobes from the convolved window function are suppressed by more than ten orders of magnitude relative to that of a nominal {\it Blackman-Harris} window for only a slight loss of resolution.}
  \label{fig:window-functions}
\end{figure}

In HERA-19 array layout, there are 30 unique baseline vectors and 8 unique baseline lengths. Fig.~\ref{fig:asm-dps-beam-chromaticity-baselines} shows the delay power spectra of foregrounds on the 8 unique baseline lengths obtained with the aforementioned models for power pattern at arbitrary local sidereal times (LST). In all these panels, the full-band spectrally weighted foreground delay power spectra obtained with achromatic, {\it Airy} and fully chromatic simulated beam patterns are shown in black, red, and blue respectively. The brightening of foreground power near the horizon limits (vertical dotted lines) due to the {\it pitchfork} effect \citep{thy15a,thy15b} is prominently seen in all cases. A clear trend of broadening of spillover-wings outside the horizon limits is seen with increasing chromaticity as the beam is changed from the achromatic to the {\it Airy} to the chromatic simulated model. For instance, the spillover from foreground delay power spectrum obtained with chromatic beam pattern is restricted to $|k_\parallel| \lesssim 0.2\,h$~Mpc$^{-1}$, with the {\it Airy} pattern it is restricted to $|k_\parallel| \lesssim 0.15\,h$~Mpc$^{-1}$, while with the achromatic beam it is restricted to $|k_\parallel| \lesssim 0.12\,h$~Mpc$^{-1}$ even on longest baseline lengths. It is also noted that despite the extreme dynamic range of the employed spectral weighting function, the tail of the foreground spillover at $|k_\parallel| \gtrsim 0.2\,h$~Mpc$^{-1}$ with an {\it Airy} beam pattern is many orders of magnitude higher than that using an achromatic beam while that from the simulated chromatic beam is even higher than from an {\it Airy} pattern by a few orders of magnitude. This signals that the foreground spillover shown are truly limited by intrinsic spectral features in the antenna beam patterns.

\begin{figure*}[htb]
  \centering
  \includegraphics[width=\linewidth]{asm_foreground_eor_beam_chromaticity_fullband_bhw2.0.eps}
  \caption{Foreground delay power spectra in units of K$^2 (h^{-1}$~Mpc$)^3$ on unique baseline lengths of HERA-19 at arbitrary LST. The baseline length and orientation (anti-clockwise from East) are annotated at the top right corner of each panel. Black, red and blue curves correspond to delay power spectra obtained with achromatic, {\it Airy} and chromatic simulated antenna beams respectively. The achromatic beam has no spectral structure, the simulated chromatic beam has maximum chromaticity while a nominal {\it Airy} pattern has intermediate level of chromaticity. The {\it pitchfork} effect is clearly visible as peaks around the horizon limits. With increase in chromaticity of the antenna beam the foreground spillover beyond the {\it foreground wedge} becomes progressively worse - the extent of foreground-spillover wings beyond the horizon limits (vertical dotted lines) and the amplitude of spillover beyond $k_\parallel\gtrsim 0.2\,h$~Mpc$^{-1}$ is most severe for the chromatic simulated beam, intermediate for an {\it Airy} pattern and negligible for an achromatic beam.}
  \label{fig:asm-dps-beam-chromaticity-baselines}
\end{figure*}

@@

It must be emphasized that such a significant spillover is not bound by the horizon limits of the {\it foreground wedge} because this is caused by spectral structure in the antenna beam pattern in addition to that caused by position-dependent geometric phase in the visibility spectra. As a result delay-based complex deconvolution techniques such as CLEAN \citep{tay99,par09,par12b} that rely on smoothness of foreground spectra and only spectral window shape will not have adequate information to accurately deconvolve intrinsic supra-horizon spillover arising from the chromaticity of the antenna beam. We leave delay power spectrum estimation using {\it foreground removal} strategy that accounts for direction-dependent and beam chromaticity-dependent effects to future work and use only simple {\it foreground avoidance} approach in this study.

@@

The effect of reflections it to shift the measured foreground power to higher modes in $\tau$ (or equivalently in $k_\parallel$) and thus cause further contamination in these higher $k_\parallel$ modes which are considered sensitive for EoR H{\sc i} signal detection. Equivalently, these delay shifts introduce ripples in the spectrum. The net chromaticity in the measurements is the product of spectral structure arising out of the intrinsic nature of foreground emission, the baseline-dependent frequency structure of the fringes, the spectral features in the antenna power pattern besides any other spectral structures in the instrument unaccounted for. Equivalently, in the Fourier domain these factors have a convolving effect. In order to isolate the requirements on beam chromaticity from other sources of chromaticity, we start with an antenna beam in which those spectral structures are absent but the intrinsic chromaticity of foregrounds and the baseline fringes are included.

>>>>

\shorttitle{Effects of Beam Chromaticity on EoR H{\sc i} Power Spectra Measurements}
\shortauthors{Thyagarajan et~al.}

@@

Unaccounted systematics from foregrounds and instruments prevent present day experiments from detecting redshifted 21~cm signal from the Epoch of Reionization (EoR). In-depth understanding of these effects is critical for such experiments to succeed. We identify spectral structures (chromaticity) in the antenna power pattern as one such systematic that significantly affects all EoR measurements. Increase in chromaticity of the antenna beam causes the already bright foreground power to leak well beyond ideal limits and contaminate the otherwise clean EoR signal modes. A primary source of this chromaticity is reflections in the antenna-feed assembly and between structures in neighboring antennas. Using precise foreground simulations taking wide-field effects into account, we provide a framework to set cosmologically-motivated upper limits on these reflections to prevent further EoR signal degradation. We use the Hydrogen Epoch of Reionization Array (HERA) as an example. Using a simple and yet high dynamic range spectral weighting scheme, we demonstrate that EoR detection using HERA will not be impeded by such spectral systematics even in a conservative scenario that does not require removal of foregrounds. We also show that all baselines in a 19-element HERA layout are capable of detecting EoR over a substantial observing window on the sky.

@@

The period in the history of the Universe characterized by the transition of neutral hydrogen in the intergalactic medium (IGM) to a fully ionized state due to the formation of radiating objects such as the first stars and galaxies is referred to as the Epoch of Reionization (EoR). This is an important period of nonlinear growth of matter density perturbations and astrophysical evolution leading to the large scale structure observed currently in the Universe. And yet, this period in the Universe's history has remained poorly probed to date.

The redshifted neutral hydrogen from the IGM in this epoch has been identified to be one of the most promising and direct probes of the EoR \citep{sun72,sco90,mad97,toz00,ili02}. Numerous experiments using low frequency radio telescopes targeting the redshifted 21~cm line from the spin-flip transition of H{\sc i} have become operational such as the Murchison Widefield Array \citep[MWA;][]{lon09,bow13,tin13}, the Donald~C.~Backer Precision Array for Probing the Epoch of Reionization \citep[PAPER;][]{par10}, the Low Frequency Array \citep[LOFAR;][]{van13} and the Giant Metrewave Radio Telescope EoR experiment \citep[GMRT;][]{pac13}. Many of these instruments have sufficient sensitivity for a statistical detection of the EoR signal by estimating the spatial power spectrum of the redshifted H{\sc i} spin temperature fluctuations \citep{bea13,thy13}. These instruments are intended to be precursors and pathfinders to the next generation of low frequency radio observatories such as the Hydrogen Epoch of Reionization Array\footnote{\url{http://reionization.org/}} \citep[HERA;][]{deb16} and the Square Kilometre Array\footnote{\url{https://www.skatelescope.org/}} (SKA). These next-generation instruments will advance the capability from a mere statistical detection of the signal to a direct three-dimensional tomographic imaging of H{\sc i} from the EoR.

@@

When expressed in the coordinate system of power spectrum measurements described by the three-dimensional wavenumber ($k$), the foreground emission is restricted to a wedge-shaped region commonly referred to as the {\it foreground wedge} \citep{bow09,liu09,liu14a,liu14b,dat10,liu11,gho12,mor12,par12b,tro12,ved12,dil13,pob13,thy13,dil14}, whereas the EoR signal has spherical symmetry due to its isotropy which appears elongated along line of sight $k$ modes due to peculiar velocity effects when dominated by matter density perturbations during early stages of reionization.

However, despite attempts to exploit this knowledge about differences in spectral and spatial symmetry in lengthy observations, analysis from the first-generation experiments has at best only led to weak constraints on reionization models \citep{ali15,dil15,pob15}. This is because their sensitivity is primarily limited by instrumental and foreground systematics. The extreme dynamic range required to suppress foreground and instrument systematics demands high precision modeling of foregrounds as observed by modern wide-field instruments. One such study was undertaken by \citet{thy15a,thy15b}. Their studies of effects of wide-field measurements of EoR power spectra have demonstrated the {\it pitchfork} effect wherein foreshortening of baselines causes a prominent enhancement of foreground power near the horizon limits of the {\it foreground wedge} as well as in the contamination beyond. It highlighted the importance of controlling the response of wide-field instruments near the horizon. The need for precise subtraction of foregrounds in wide-field measurements was also highlighted in \citet{pob16}.

In this paper, we explore yet another phenomenon that inherently extends foreground power beyond the {\it wedge}, namely, that arising from the chromaticity of the antenna power pattern. While highlighting the importance of such instrumental systematics on future EoR experiments, we demonstrate that the current HERA instrument design will exceed these specifications to detect the EoR signal with high significance. This will hold when data is limited by foregrounds and systematics even in the most conservative scenario which involves neither foreground subtraction nor optimal analysis techniques but only relies on a simple spectral weighting scheme.

This paper is organized as follows. \S\ref{sec:HERA} introduces the HERA instrument. A brief summary of the delay spectrum technique used extensively in this analysis and the recently confirmed {\it pitchfork} effect are presented in \S\ref{sec:delay-spectrum}. \S\ref{sec:sim} describes foreground simulations including antenna beam pattern, all-sky foreground models and two independent EoR models. \S\ref{sec:beam-chromaticity} investigates the effects of chromaticity of antenna beam on the resulting delay power spectrum and the cosmologically motivated constraints it places on reflections in the instrument. Lower limits on sensitivity of HERA under simple {\it foreground avoidance} criteria are demonstrated in \S\ref{sec:eor-sensitivity}. Our findings are summarized in \S\ref{sec:summary}.

@@

The delay spectrum has a close resemblance to cosmological H~{\sc i} spatial power spectrum and is defined as:
\begin{align}\label{eqn:delay-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{1}{\Omega\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \label{eqn:kperp-baseline}\\
  k_\parallel &\equiv \frac{2\pi\tau\,f_{21}H_0\,E(z)}{c(1+z)^2}, \label{eqn:kprll-delay}
\end{align}
where, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H~{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology, and following \citet{par14},
\begin{align}
  \Omega\Delta B &= \iint \left|A(\hat{\boldsymbol{s}},f)\,W(f)\right|^2\,\dif\Omega\,\dif f.
\end{align}

@@

We note that these visibilities and delay power spectra are dependent on right ascension (RA) of the pointing as well as the frequency band. However, we defer expressing them explicitly as a function of pointing and center frequency (or redshift) to later sections where it becomes necessary.

@@

It was also demonstrated in these studies that design of antenna power pattern, specifically its amplitude near the horizon, is an important tool in mitigating foreground contamination caused by these wide-field effects. A dish characterized by a nominal {\it Airy} pattern was found to mitigate this contamination by over four and two orders of magnitude relative to a dipole and a phased array of dipoles respectively. HERA has significantly based its antenna design principles on these findings in choosing its antenna geometry.

In this paper, we investigate the spectral properties of the proposed dish design through their effects on the resultant foreground delay power spectrum and the constraints they place on attenuation required to suppress reflections between the antenna-feed assembly and antenna pairs to within tolerable limits.

@@

In a related series of papers, \citet{neb16} discuss the agreement of these simulated antenna beam patterns with actual measurements and \citet{ewa16} model the reflections and return loss expected in the proposed antenna-feed assembly. Our focus in this paper is to investigate chromaticity of antenna power patterns from the point of view of their impact on foreground contamination in delay power spectra.

@@

For reference, we use two models of EoR. In the first, simulations of the H{\sc i} signal were created using the publically available 21cmFAST\footnote{\url{http://homepage.sns.it/mesinger/DexM\_\_\_21cmFAST.html}} code described in \citet{mes11}. The code uses the excursion set formalism of \citet{fur04a} to generate ionization and 21cm brightness fields for numerous redshifts. The model shown in this paper assumes the same fiducial values as in \citet{ewa16}: $T_\text{vir}^\text{min} = 2 \times 10^4$~K (virial temperature of minimum mass of dark matter halos that host ionizing sources), $\zeta = 20$ (ionization efficiency), and $R_\text{mfp}=15$~Mpc (mean free path of UV photons) which predicts the redshift of 50\% ionization (and hence a peak in the power spectrum signal) to be at $z=8.5$. The second EoR model is borrowed from the simulations described in detail in \citet{lid08}. Hereafter, we refer to these two as EoR models 1 and 2 respectively.

@@

The equation for delay spectrum describes the mapping between sky location of a foreground object and delay. The chromatic nature (variation with frequency) of the antenna power pattern results in a convolution of the geometrical mapping with the delay response of spectral chromaticity of the power pattern. We discuss the resulting spillover of foreground power beyond the horizon delay limits.

@@

Since our aim in this paper is to quantify foreground contamination and EoR sensitivity without employing any sophisticated {\it foreground removal} techniques, we turn out attention to {\it foreground avoidance} strategy that employs spectral weighting technique. In such a scheme, specific choices for spectral weighting function, $W(f)$, have been found to be effective in reducing foreground contamination by many orders of magnitude \citep{thy13} and is regularly used in EoR data analysis \citep{par12a}. For instance, a {\it Blackman-Harris} function \citep{har78} has a dynamic range of $\sim$100 -- 120~dB in delay power spectrum and a reduced effective bandwidth, with:
\begin{align}\label{eqn:Beff}
  \epsilon &= \frac{1}{B} \int_{-B/2}^{+B/2} |W(f)|^2\,\dif f, \\
  \textrm{and,}\quad B_\textrm{eff} &= \epsilon\,B,
  % B_\textrm{eff} &= \int_{-B/2}^{+B/2} |W(f)|^2\,\dif f.
\end{align}
where, $1-\epsilon$ is the loss in overall spectral sensitivity and $B_\textrm{eff}$ is the effective bandwidth. For a {\it Blackman-Harris} window function denoted by $W_\textrm{BH}(f)$, $\epsilon \approx 50\%$. % Hence, the window spans a total of $2N$ frequency channels each of width $\Delta f$ to have $B_\textrm{eff} \approx N\,\Delta f$.

Considering that the dynamic range required to suppress foreground contamination in the {\it EoR window} may probably be even higher than that provided by a {\it Blackman-Harris} window function, we use a modified version given by the convolution of a {\it Blackman-Harris} window with itself:
\begin{align}\label{eqn:bhw2}
  W(f) &= W_\textrm{BH}(f) \ast W_\textrm{BH}(f).
\end{align}

Fig.~\ref{fig:window-functions} shows the {\it Blackman-Harris} window (gray) and  the window in equation~\ref{eqn:bhw2} (black) in frequency (top) and their power responses in delay (bottom). In its Fourier domain, this window has a response obtained by squaring the {\it Blackman-Harris} window response and thus increases the dynamic range further in the power response by another 100 -- 120~dB with $\epsilon\approx 42\%$.

\begin{figure}[htb]
  \centering
  \includegraphics[width=\linewidth]{window_function_modifications.eps}
  \caption{Choices for spectral weighting functions, $W(f)$ ({\it top}) and their delay power spectrum responses, $|\widetilde{W}(\tau)|^2$ ({\it bottom}). The gray curves correspond to a {\it Blackman-Harris} window while those in black correspond to a {\it Blackman-Harris} window convolved with itself. The overall sensitivity of the former is $\epsilon \approx 50\%$ of a rectangular window while that of the latter is $\epsilon \approx 42\%$. As a result, the main lobe of the convolved window function response is slightly wider than that of a {\it Blackman-Harris} window. However, the sidelobes from the convolved window function are suppressed by more than ten orders of magnitude relative to that of a nominal {\it Blackman-Harris} window.}
  \label{fig:window-functions}
\end{figure}

This is an enormous gain in sidelobe characteristics for a negligible loss of sensitivity relative to a simple {\it Blackman-Harris} weighting. This will ensure that the contamination resulting from spillover of foregrounds along the line-of-sight $k$-modes are limited only by intrinsic spectral structures in the foregrounds and the beam patterns and not by sidelobes from spectral weighting. In this paper, we apply this modified spectral weighting function defined in equation~\ref{eqn:bhw2} whenever foregrounds are represented in the Fourier delay domain.

In HERA-19 array layout, there are 30 unique baseline vectors and 8 unique baseline lengths. Fig.~\ref{fig:asm-dps-beam-chromaticity-baselines} shows the delay power spectra of foregrounds on the 8 unique baseline lengths obtained with the aforementioned models for power pattern at arbitrary pointings on the sky. In all these panels, the full-band spectrally weighted foreground delay power spectra obtained with achromatic, {\it Airy} and fully chromatic simulated beam patterns are shown in black, red, and blue respectively. The brightening of foreground power near the horizon limits (vertical dotted lines) due to the {\it pitchfork} effect \citep{thy15a,thy15b} is prominently seen in all cases. A clear trend of broadening of spillover-wings outside the horizon limits is seen with increasing chromaticity as the beam is changed from the achromatic to the {\it Airy} to the chromatic simulated model. For instance, the spillover from foreground delay power spectrum obtained with chromatic beam pattern is restricted to $|k_\parallel| \lesssim 0.2\,h$~Mpc$^{-1}$, with the {\it Airy} pattern it is restricted to $|k_\parallel| \lesssim 0.15\,h$~Mpc$^{-1}$, while with the achromatic beam it is restricted to $|k_\parallel| \lesssim 0.12\,h$~Mpc$^{-1}$ even on longest baseline lengths. It is also noted that despite the extreme dynamic range of the employed spectral weighting function, the tail of the foreground spillover at $|k_\parallel| \gtrsim 0.2\,h$~Mpc$^{-1}$ with an {\it Airy} beam pattern is many orders of magnitude higher than that using an achromatic beam while that from the simulated chromatic beam is even higher than from an {\it Airy} pattern by a few orders of magnitude. This signals that the foreground spillover shown are truly limited by intrinsic spectral features in the antenna beam patterns.

\begin{figure*}[htb]
  \centering
  \includegraphics[width=\linewidth]{asm_foreground_eor_beam_chromaticity_fullband_bhw2.0.eps}
  \caption{Foreground delay power spectra in units of K$^2 (h^{-1}$~Mpc$)^3$ on unique baseline lengths of HERA-19 at arbitrary sky pointings. The baseline length and orientation (anti-clockwise from East) are annotated at the top right corner of each panel. Black, red and blue curves correspond to delay power spectra obtained with achromatic, {\it Airy} and chromatic simulated antenna beams respectively. The achromatic beam has no spectral structure, the simulated chromatic beam has maximum chromaticity while a nominal {\it Airy} pattern has intermediate level of chromaticity. The {\it pitchfork} effect is clearly visible as peaks around the horizon limits. With increase in chromaticity of the antenna beam the foreground spillover beyond the {\it foreground wedge} becomes progressively worse - the extent of foreground-spillover wings beyond the horizon limits (vertical dotted lines) and the amplitude of spillover beyond $k_\parallel\gtrsim 0.2\,h$~Mpc$^{-1}$ is most severe for the chromatic simulated beam, intermediate for an {\it Airy} pattern and negligible for an achromatic beam.}
  \label{fig:asm-dps-beam-chromaticity-baselines}
\end{figure*}

@@

It must be emphasized that such a significant spillover is not bound by the horizon limits of the {\it foreground wedge} because this is caused by spectral structure in the antenna beam pattern which is independent of the position-dependent geometric delays. As a result delay-based complex deconvolution techniques such as CLEAN \citep{tay99,par09,par12b} that rely on smoothness of foreground spectra and only spectral window shape will not have adequate information to accurately deconvolve intrinsic supra-horizon spillover arising from the chromaticity of the antenna beam. We leave delay power spectrum estimation using {\it foreground removal} strategy that accounts for direction-dependent and beam chromaticity-dependent effects to future work.

@@

The effect of reflections it to shift the measured foreground power to higher modes in $\tau$ (or equivalently in $k_\parallel$) and thus cause further contamination in these higher $k_\parallel$-modes which are considered sensitive for EoR H{\sc i} signal detection. Equivalently, these delay shifts introduce ripples in the spectrum. The net chromaticity in the measurements is the product of spectral structure arising out of the intrinsic nature of foreground emission, the baseline- and position-dependent frequency structure of the fringes, the spectral features in the antenna power pattern besides any other spectral structures in the instrument unaccounted for. Equivalently, in the Fourier domain these factors have a convolving effect. In order to isolate the requirements on beam chromaticity from other sources of chromaticity, we start with an antenna beam in which those spectral structures are absent but the intrinsic chromaticity of foregrounds and the baseline fringes are included.

We define the required attenuation on the reflected foreground power as the ratio: 
\begin{align}\label{eqn:attenuation}
  \Gamma_{k_\textrm{p}}(\tau) \geq \min_{|k_\parallel|>k_\textrm{p}}\left\{\frac{P_\textrm{FG}(k_\parallel - \frac{\dif k_\parallel}{\dif \tau}\,\tau)}{P_\textrm{H{\sc i}}(k_\parallel)}\right\}^{1/2},
\end{align}
where, $P_\textrm{H{\sc i}}(k_\parallel)$ is the EoR H{\sc i} power spectrum for the chosen antenna spacing, $P_\textrm{FG}(k_\parallel)$ is the foreground delay power spectrum obtained with the spectral weighting function applied over the full band of the three uniquely oriented 14.6~m antenna spacings further averaged over a 0--12 hr local sidereal time (LST) range and over both positive and negative $k_\parallel$-modes, $\tau$ is the delay caused by reflections, and $\dif k_\parallel/\dif \tau$ is the {\it jacobian} in the transformation of $\tau$ to $k_\parallel$ (see equation~\ref{eqn:kprll-delay}). Equivalently, $\Gamma_{k_\textrm{p}}(\tau)$ is determined by the requirement that the reflected foreground power obtained by shifting in delay is below the EoR H{\sc i} signal power in line-of-sight spatial scales of interest given by $|k_\parallel|>k_\textrm{p}$. For this analysis constraining reflections, we use EoR model 1 at 150~MHz ($z\approx 8.47$).
